---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


# Loading data
```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
  unzip('activity.zip')
}
activityData <- read.csv('activity.csv')
```

#Calculate the total number of steps taken per day
```{r}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```

#Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
qplot(stepsByDay, xlab='Number of steps per day', ylab='Frequency',binwidth=500)
```

#Mean and median number of steps taken each day
```{r}
stepsByDay_mean <- mean(stepsByDay)
stepsByDay_median <- median(stepsByDay)
```

* Mean: `r stepsByDay_mean` 
* Median: `r stepsByDay_median`


#Time series plot of the average number of steps taken
```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps))+geom_line()+xlab('Interval')+ylab('average number of steps')
```

#The 5-minute interval that, on average, contains the maximum number of steps
```{r}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```
* Most Steps at: `r timeMostSteps`


#Code to describe and show a strategy for imputing missing data
```{r}
library(Hmisc)
numMissingValues <- length(which(is.na(activityData$steps)))
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```

* Number of missing values: `r numMissingValues`

#Histogram of the total number of steps taken each day after missing values are imputed
```{r}
stepsByDay_imputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum, na.rm=TRUE)
qplot(stepsByDay_imputed, xlab='Number of steps per day (Imputed)', ylab='Frequency',binwidth=500)
```


#Mean and median number of steps taken each day (Imputed)
```{r}
stepsByDay_mean_imputed <- mean(stepsByDay_imputed)
stepsByDay_median_imputed <- median(stepsByDay_imputed)
```
* Mean (Imputed): `r stepsByDay_mean_imputed`
* Median (Imputed):  `r stepsByDay_median_imputed`


#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + geom_line() + facet_grid(dateType ~ .) +xlab("5-minute interval") + ylab("avarage number of steps")
```
